---
title: "Fitting the the two-parameter Weibull Distribution"
# author: "Fatih Kızılaslan"
# date: "`r format(Sys.time(), '%B %d, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting the two-parameter Weibull Distribution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```


## Introduction

This vignette presents the `fitWD()` function, which is used to estimate the parameters of the two-parameter Weibull Distribution (WD) using several estimation methods, including maximum likelihood (ML), least squares (LS), weighted least squares (WLS), and maximum product of spacings (MPS).

A random variable $X$ is said to follow a two-parameter WD its cumulative distribution function (CDF) and probability density function (PDF) are given by
$$
    F(x) = 1- e^{-a x^b},
$$
and
$$
    f(x) = a b x^{b-1} e^{-a x^b},
$$    
where $x>0$, $a>0$ is the scale parameter and $b > 0$ is a shape parameter.

### Estimation methods

Let a random sample from $X \sim WD(a, b)$ are observed as $x_i, \; i=1,\dots, n$.
For all estimation methods, optimization is performed using `stats::optim`.

- **Maximum Likelihood Estimation (MLE)**

The MLEs are obtained by maximizing the log-likelihood function:
$$
\ell(a,b; \mathbf{x}) = \sum_{i=1}^{n} \log \big( a b x_i^{b-1} \big) -a \sum_{i=1}^{n} x_i^b.
$$ 

- **Least Squares Estimation (LSE)**

The LSE is obtained by minimizing the squared differences between the theoretical and empirical CDF values at the ordered sample points.

Let $x_{(1)}<x_{(2)}< \dots < x_{(n)}$ denote the ordered samples. The objective function is:
$$
Q(a,b) = \sum_{i=1}^n \bigg( F(x_{(i)};a,b) - \frac{i-0.3}{n+0.4}   \bigg)^2.
$$

- **Weighted Least Squares Estimation (WLSE)**

The WLSE extends LSE by introducing weights:
$$
Q^W(a,b) = \sum_{i=1}^n w_i \bigg( F(x_{(i)};a,b) - \frac{i-0.3}{n+0.4}   \bigg)^2,
$$
where $w_i = \frac{(n+1)^2(n+2)}{i(n-i+1)}, \; i=1,\dots,n.$

- **Maximum Product of Spacings (MPS)** 

The MPS estimates are obtained by maximizing:
$$
    M(a,b) = \sum_{i=1}^{n+1} \frac{1}{n+1} \log \big( F(x_{(i)};a,b) - F(x_{(i-1)};a,b) \big),
$$
where $F(x_{(0)};a,b) = 0$ and $F(x_{(n+1)};a,b) = 1$.

## Using `fitWD()`

To illustrate the use of `fitWD()`, we consider a simulated example.

```{r}
rm(list = ls())

# Load the package
library(SSReliabilityClaytonMWD)

# generate data from MWD(a, b, lambda) with lambda = 0 reduces two-parameter Weibull distribution
n <- 50
a <- 0.75; b <- 1.25; lambda <- 0
set.seed(123)
dat <- rMweibull(n, a, b, lambda)
# random initial points 
init <- runif(2)
```

MLE of the parameters:
```{r}
# Fit WD to `dat`
fit.mle <- fitWD(data = dat, est.method = "mle", opt.method = "L-BFGS-B", starts = init,
                  lower = c(1e-05,1e-05), upper = c(Inf,Inf), hessian = FALSE )
fit.mle$estimates
```

LSE of the parameters:
```{r}
fit.lse <- fitWD(data = dat, est.method = "lse", opt.method = "L-BFGS-B", starts = init,
                   lower = c(1e-05,1e-05), upper = c(Inf,Inf), hessian = FALSE )
fit.lse$estimates
```

WLSE of the parameters:
```{r}
fit.wlse <- fitWD(data = dat, est.method = "wlse", opt.method = "L-BFGS-B", starts = init,
                    lower = c(1e-05,1e-05), upper = c(Inf,Inf), hessian = FALSE )
fit.wlse$estimates
```

MPS of the parameters:
```{r}
fit.mps <- fitWD(data = dat, est.method = "mps", opt.method = "L-BFGS-B", starts = init,
                   lower = c(1e-05,1e-05), upper = c(Inf,Inf), hessian = FALSE )
fit.mps$estimates
```

